```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(targets)
```

# Debugging {#debugging}

If one of your targets fails, first look up the error message in `tar_meta()`. If that does not help, try one of the following techniques.

## Workspaces

Workspaces are special lightweight reference files that allow `tar_workspace()` to recreate the runtime environment of a target. This lets you troubleshoot issues outside the pipeline in an interactive session. There are two ways to save a workspace file:

1. Set `error = "workspace"` in `tar_option_set()` or `tar_target()`. Then, `tar_make()` and friends will save a workspace file for every target that errors out.
1. In the `workspaces` argument of `tar_option_set()`, specify the targets for which you want to save workspaces. Then, run `tar_make()` or similar. A workspace file will be saved for each existing target, regardless of whether the target runs or gets skipped in the pipeline.

Here is an example of (1).

```{r, eval = FALSE}
# _targets.R file:
options(tidyverse.quiet = TRUE)
library(targets)
library(tidyverse)
options(crayon.enabled = FALSE)
tar_option_set(error = "workspace")
f <- function(arg, value) {
  stopifnot(arg < 4)
}
list(
  tar_target(x, seq_len(4)),
  tar_target(
    y,
    f(arg = x, value = "succeeded", a = 1, b = 2, key = "my_api_key"),
    pattern = map(x) # The branching chapter describes patterns.
  )
)
```

```{r, eval = FALSE}
# R console:
tar_make()
#> ● run target x
#> ● run branch y_29239c8a
#> ● run branch y_7cc32924
#> ● run branch y_bd602d50
#> ● run branch y_05f206d7
#> x error branch y_05f206d7
#> ● save workspace y_05f206d7
#> Error : x < 4 is not TRUE .
#> Error: callr subprocess failed: x < 4 is not TRUE .
```

One of the `y_*******` targets errored out.

```{r, eval = FALSE}
failed <- tar_meta(fields = error) %>%
  na.omit() %>%
  pull(name)

print(failed)
#> [1] "y_05f206d7"
```

`tar_workspace()` reads the special metadata in the workspace file and then loads the target's dependencies from various locations in `_targets/objects` and/or the [cloud](#cloud). It also sets the random number generator seed to the seed of the target, loads the required packages, and runs `_targets.R` to load other global object dependencies such as functions. 

```{r, eval = FALSE}
tar_workspace(y_05f206d7)
```

We now have the dependencies of `y_05f206d7` in memory, which allows you to try out any failed function calls in your local R session. ^[In addition, current random number generator seed (`.Random.seed`) is also the value `y_05f206d7` started with.] ^[When you are finished debugging, you can remove all workspace files with `tar_destroy(destroy = "workspaces")`.]

```{r, eval = FALSE}
print(x)
#> [1] 4
f(arg = 0, value = "my_value", a = 1, b = 2, key = "my_api_key")
#> [1] "my_value"
f(arg = x, value = "my_value", a = 1, b = 2, key = "my_api_key")
#> Error in f(x) : x < 4 is not TRUE
```

Keep in mind that that although the dependencies of `y_05f206d7` are in memory, the arguments of `f()` are not.

```{r, eval = FALSE}
arg
#> Error: object 'arg' not found
value
#> Error: object 'value' not found
```

## Interactive debugging

Interactive debugging offers a [`browser()`](https://adv-r.hadley.nz/debugging.html#browser)-like debugging experience. Unlike workspaces, which most users access after the pipline finishes, interactive debugging lets you step through a target's command while the target is running. This approach is not feasible if the pipeline is running in a non-interactive process or requires resources that are only available on a remote compute node of a cluster. However, when possible, it may be more convenient than workspaces. In our example above, `tar_workspace()` automatically loaded a branch of the dependency `x`, but the user still needed to manually supply the rest of the arguments of `f()`. During interactive debugging, you will instead be able to call `debug(f)` and then press `c` to immediately drop into the function environment where the remaining variables are already defined.

Follow these steps to use interactive debugging.

1. In `_targets.R`, write a call to `tar_option_set()` with `debug` equal to the target name. Consider also setting `cue` equal to `tar_cue(mode = "never")` so `tar_make()` reaches the target you want to debug more quickly.
1. Launch a fresh clean new interactive R session with the `_targets.R` script in your working directory.
1. Run `targets::tar_make()` (or `targets::tar_make_clustermq()`, or `targets::tar_make_future()`) with `callr_function = NULL`.
1. When `targets` reaches the target you selected to debug, your R session will start an interactive debugger, and you should see `Browse[1]>` in your console. Run `targets::tar_name()` to verify that you are debugging the correct target.
1. Interactively run any R code that helps you troubleshoot the problem. For example, if the target invokes a function `f()`, enter `debug(f)` and then `c` to immediately enter the function's calling environment where all its arguments are defined.

To try it out yourself, write the following `_targets.R` file.

```{r, eval = FALSE}
# _targets.R
library(targets)
tar_option_set(debug = "b")
f <- function(x, another_arg = 123) x + another_arg
list(
  tar_target(a, 1),
  tar_target(b, f(a))
)
```

Then, call `tar_make(callr_function = NULL)` to drop into a debugger at the command of `b`. 

```{r, eval = FALSE}
# R console
tar_make(callr_function = NULL)
#> ● run target a
#> ● run target b
#> Called from: eval(expr, envir)
Browse[1]>
```

When the debugger launches, run `targets::tar_name()` to confirm you are running the correct target.

```{r, eval = FALSE}
Browse[1]> targets::tar_name()
#> [1] "b"
```

In the debugger, the dependency targets of `b` are available in the current environment, and the global objects and functions are available in the parent environment.

```{r, eval = FALSE}
Browse[1]> ls()
#> [1] "a"
Browse[1]> a
#> [1] 1
Browse[1]> ls(parent.env(environment()))
#> [1] "f"
Browse[1]> f(1)
#> [1] 124
```

Enter `debug(f)` to debug the function `f()`, and press `c` to enter the function's calling environment where `another_arg` is defined.

```{r, eval = FALSE}
Browse[1]> debug(f)
Browse[1]> c
#> debugging in: f(a)
#> debug at _targets.R#3: x + another_arg
Browse[2]> ls()
#> [1] "another_arg" "x"   
Browse[2]> another_arg
#> [1] 123
```

Visit [this page](https://rstats.wtf/debugging-r-code.html) for more information on debugging R code.
