---
execute:
  freeze: auto
---

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = TRUE)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(targets)
```

# Debugging {#debugging}

This chapter describes how to identify and solve errors in `targets` pipelines.

## Challenges

Pipelines are automated for reproducibility. When you run `tar_make()`, the following steps occur:

1. `targets` starts a new non-interactive R process using `callr::r()`.[^callr]
1. The non-interactive R process runs `_targets.R` to load functions and objects, set options, and read the pipeline definition.
1. The non-interactive R process orchestrates the targets, running or skipping each as necessary.

[^callr]: The `callr_function` argument controls which [`callr`](https://github.com/r-lib/callr) function initializes the process.

Your interactive R session cannot access this non-interactive R process. This makes pipelines more difficult to debug than the informal commands you may type into the R console or the [RStudio IDE](https://www.rstudio.com/products/rstudio/).

## Example

The following pipeline simulates a repeated measures dataset and analyzes it with generalized least squares.

```{r, eval = FALSE}
# _targets.R
library(targets)
tar_option_set(
  packages = c("broom", "broom.mixed", "dplyr", "nlme", "tibble", "tidyr")
)

simulate_data <- function(units) {
  tibble(unit = seq_len(units), factor = rnorm(n = units, mean = 3)) %>%
    expand_grid(measurement = seq_len(4)) %>%
    mutate(outcome = sqrt(factor) + rnorm(n()))
}

analyze_data <- function(data) {
  gls(
    model = outcome ~ factor,
    data = data,
    correlation = corSymm(form = ~ measurement | unit),
    weights = varIdent(form = ~ 1 | measurement)
  ) %>%
    tidy(conf.int = TRUE, conf.level = 0.95)
}

list(
  tar_target(name = dataset1, command = simulate_data(100)),
  tar_target(name = model, command = analyze_data(dataset1))
)
```

This pipeline has an error.

```{r, eval = FALSE}
# R console
tar_make()
#> • start target dataset1
#> • built target dataset1 [0.499 seconds]
#> • start target model
#> ✖ error target model
#> • end pipeline [0.631 seconds]
#> Warning messages:
#> 1: NaNs produced 
#> 2: 1 targets produced warnings. Run tar_meta(fields = warnings, complete_only = TRUE) for the messages. 
#> Error:
#> ! Error running targets::tar_make()
#>   Target errors: targets::tar_meta(fields = error, complete_only = TRUE)
#>   Tips: https://books.ropensci.org/targets/debugging.html
#>   Last error: missing values in object
```

## Finish the pipeline anyway

Even if you hit an error, you can still finish the successful parts of the pipeline. The `error` argument of `tar_option_set()` and `tar_target()` tells each target what to do it hits an error. For example, `tar_option_set(error = "null")` tells errored targets to return `NULL`. The output as a whole will not be correct or up to date, but the pipeline will finish so you can look at preliminary results. This is especially helpful with [dynamic branching](#dynamic).

```{r, eval = FALSE}
# _targets.R
library(targets)
tar_option_set(
  packages = c("broom", "broom.mixed", "dplyr", "nlme", "tibble", "tidyr")
  error = "null" # produce a result even if the target errored out.
)

# Functions etc...
```

```{r, eval = FALSE}
# R console
tar_make()
#> • start target dataset1
#> • built target dataset1 [0.657 seconds]
#> • start target model
#> ✖ error target model
#> ✖ error target model
#> • end pipeline [0.783 seconds]
#> Warning messages:
#> 1: NaNs produced 
#> 2: 1 targets produced warnings. Run tar_meta(fields = warnings, complete_only = TRUE) for the messages.

# We do have a result for target {model}.
tar_read(model)
#> NULL

# But it is not up to date.
tar_outdated()
#> [1] "model"
```


## Error messages

Still, it is important to fix known errors. The metadata in `_targets/meta/meta` is a good place to start. It stores the most recent error and warning messages for each target. `tar_meta()` can retrieve these messages.

```{r, eval = FALSE}
# R console
tar_meta(fields = error, complete_only = TRUE)
#> # A tibble: 1 × 2
#>   name  error                   
#>   <chr> <chr>                   
#> 1 model missing values in object
```

```{r, eval = FALSE}
# R console
tar_meta(fields = warnings, complete_only = TRUE)
#> # A tibble: 1 × 2
#>   name     warnings     
#>   <chr>    <chr>        
#> 1 dataset1 NaNs produced
```

It looks like missing values in the data are responsible for the error in the `model`. Maybe this clue alone is enough to repair the code.[^missing] If not, read on.

[^missing]: You can fix the bug by either removing the missing values from the dataset or by setting `na.action = na.omit` in `gls()`.

## Functions

Most errors are come from custom user-defined functions like `simulate_data()` and `analyze_data()`. See if you can reproduce the error in the R console.

```{r, eval = FALSE}
# R console
library(targets)

# Restart your R session.
rstudioapi::restartSession()

# Loads globals like tar_option_set() packages, simulate_data(), and analyze_data():
tar_load_globals()

# Load the data that the target depends on.
tar_load(dataset1)

# Run the command of the errored target.
analyze_data(dataset1)
#> Error in na.fail.default(list(measurement = c(1L, 2L, 3L, 4L, 1L, 2L,  : 
#>   missing values in object
```

If you see the same error here that you saw in the pipeline, then good! Now that you are in an interactive R session, all the [usual debugging techniques and tools](https://adv-r.hadley.nz/debugging.html) such as `debug()` and `browser()` can help you figure out how to fix your code, and you can exclude `targets` from the rest of the debugging process.

```{r, eval = FALSE}
# R console
debug(analyze_data)
analyze_data(dataset1)
#> debugging in: analyze_data(dataset)
#> ...
Browse[2]> anyNA(dataset$outcome) # Do I need to handle missing values?
#> [1] TRUE
```

In some cases, however, you may not see the original error:

```{r, eval = FALSE}
# R console
dataset <- simulate_data(100)
analyze_data(dataset)
#> # A tibble: 2 × 7
#>   term        estimate std.error statistic  p.value conf.low conf.high
#> * <chr>          <dbl>     <dbl>     <dbl>    <dbl>    <dbl>     <dbl>
#> 1 (Intercept)    0.595    0.152       3.92 1.05e- 4    0.297     0.892
#> 2 factor         0.362    0.0476      7.61 2.04e-13    0.269     0.455
```

Above, the random number generator seed in your local session is different from the seed assigned to the target in the pipeline. The dataset from the pipeline has missing values, whereas the one in the local session does not.

If you cannot reproduce the error in an interactive R session, read on.

## System issues

If you see an error in the pipeline but not your local interactive R session, then the bug could be system-related. For example, [this issue](https://github.com/ropensci/targets/discussions/506) was originally reported as a bug in `targets` but actually turned out to be a bug in the [interaction between packages `renv` and `callr`](https://github.com/rstudio/renv/issues/773#issuecomment-973765196). (Remember, `targets` uses `callr::r()` to launch the pipeline). To check for issues like that one, try running `tar_make()` with `callr_function = NULL` to avoid `callr` entirely:

```{r, eval = FALSE}
# R console
rstudioapi::restartSession() # Remove in-memory detritus.
targets::tar_make(callr_function = NULL)
```

In addition, try running your code with `callr` outside the pipeline:

```{r, eval = FALSE}
# R console
callr::r(
  func = function() {
    library(targets)
    tar_load_globals()
    dataset <- simulate_data(100)
    analyze_data(dataset)
  },
  show = TRUE
)
```

Similarly, you can isolate many [high-performance computing](https://books.ropensci.org/targets/hpc.html) problems by directly invoking [`clustermq`](https://github.com/mschubert/clustermq) or [`future`](https://github.com/HenrikBengtsson/future). Examples:

```{r, eval = FALSE}
# R console with clustermq:
# see https://mschubert.github.io/clustermq/articles/userguide.html
options(clustermq.scheduler = "multiprocess")
clustermq::Q(
  function(arg) {
    library(targets)
    tar_load_globals()
    dataset <- simulate_data(100)
    analyze_data(dataset)
  },
  arg = 1,
  n_jobs = 1
)
```

```{r, eval = FALSE}
# R console with future:
# see https://future.futureverse.org
future::plan(future::multisession)
f <- future::future(
  expr = {
    library(targets)
    tar_load_globals()
    dataset <- simulate_data(100)
    analyze_data(dataset)
  },
  seed = TRUE
)
future::value(f)
```

If you successfully reproduce the bug without using `targets`, then the problem becomes much smaller and much easier to solve. At that point, you can completely exclude `targets` from the rest of the debugging process.

## Pause the pipeline with `browser()`

Sometimes, you may still need to run the pipeline to find the problem. The following trick lets you pause the pipeline and tinker with a running target interactively:

1. Insert [`browser()`](https://adv-r.hadley.nz/debugging.html#browser) into the function that produces the error.
1. Restart your R session to remove detritus from memory.[^detritus]
1. Call `tar_make()` with `callr_function = NULL` to run the pipeline in your interactive session without launching a new `callr::r()` process.
1. Poke around until you find the bug.

[^detritus]: With `callr_function = NULL`, a messy local R environment can accidentally change the functions and objects that a target depends on, which can invalidate those targets and erase hard-earned results that were previously correct. This is why `targets` uses `callr` in the first place, and it is why `callr_function = NULL` is for debugging only. If you do need `callr_function = NULL`, please restart your R session first.

```{r, eval = FALSE}
# _targets.R
# ...
analyze_data <- function(data) {
  browser() # Pause the pipeline here.
  gls(
    model = outcome ~ factor,
    data = data,
    correlation = corSymm(form = ~ measurement | unit),
    weights = varIdent(form = ~ 1 | measurement)
  ) %>%
    tidy(conf.int = TRUE, conf.level = 0.95)
}
# ...
```

```{r, eval = FALSE}
# R console
library(targets)

# Restart your R session.
rstudioapi::restartSession()

# Run the pipeline in your interactive R session (no callr process)
tar_make(callr_function = NULL)
#> ✔ skip target dataset1
#> • start target model
#> Called from: analyze_data(dataset1)

# Tinker with the R session to see if you can reproduce the error.
Browse[1]> model <- gls(
+   model = outcome ~ factor,
+   data = data,
+   correlation = corSymm(form = ~ measurement | unit),
+   weights = varIdent(form = ~ 1 | measurement)
+ )
#> Error in na.fail.default(list(measurement = c(1L, 2L, 3L, 4L, 1L, 2L,  : 
#>   missing values in object

# Figure out what it would take to fix the error.
Browse[1]> model <- gls(
+   model = outcome ~ factor,
+   data = data,
+   correlation = corSymm(form = ~ measurement | unit),
+   weights = varIdent(form = ~ 1 | measurement),
+   na.action = na.omit
+ )

# Confirm that the bug is fixed.
Browse[1]> tidy(model, conf.int = TRUE, conf.level = 0.95)
#> # A tibble: 2 × 7
#>   term        estimate std.error statistic       p.value conf.low conf.high
#> * <chr>          <dbl>     <dbl>     <dbl>         <dbl>    <dbl>     <dbl>
#> 1 (Intercept)    0.795    0.148       5.36 0.000000145      0.504     1.09 
#> 2 factor         0.275    0.0466      5.92 0.00000000717    0.184     0.367
```

## Pause the pipeline with the `debug` option

It may be too tedious to comb through all targets with `browser()`. For example, what if the pipeline has hundreds of simulated datasets? The following pipeline simulates 100 datasets with 58 experimental units each and 100 datasets with 70 experimental units each. Each dataset is analyzed with `gls()`. `tar_map_rep()` from the `tarchetypes` package organizes this simulation structure and batches the replications for computational efficiency.

```{r, eval = FALSE}
# _targets.R
library(targets)
library(tarchetypes)

tar_option_set(
  packages = c("broom", "broom.mixed", "dplyr", "nlme", "tibble", "tidyr")
)

simulate_data <- function(units) {
  tibble(unit = seq_len(units), factor = rnorm(units, mean = 3)) %>%
    expand_grid(measurement = seq_len(4)) %>%
    mutate(outcome = sqrt(factor) + rnorm(n()))
}

analyze_data <- function(data) {
  gls(
    model = outcome ~ factor,
    data = data,
    correlation = corSymm(form = ~ measurement | unit),
    weights = varIdent(form = ~ 1 | measurement)
  ) %>%
    tidy(conf.int = TRUE, conf.level = 0.95)
}

simulate_and_analyze_one_dataset <- function(units) {
  data <- simulate_data(units)
  analyze_data(data)
}

list(
  tar_map_rep( # from the {tarchetypes} package
    name = analysis,
    command = simulate_and_analyze_one_dataset(units), # 100 rows per dataset
    values = data.frame(units = c(58, 70)), # 2 data size scenarios.
    names = all_of("units"), # which columns of values should be used to name the targets
    batches = 20, # For each scenario, divide the 100 simulations into 20 dynamic branch targets.
    reps = 5 # Each branch target (batch) runs simulate_and_analyze_one_dataset(n = 100) 5 times.
  )
)
```

```{r, eval = FALSE}
# R console
tar_make()
#> ✔ skip target analysis_batch
#> • start branch analysis_58_550d992c
#> • built branch analysis_58_550d992c [1.2 seconds]
#> • start branch analysis_58_582bca0a
#> • built branch analysis_58_582bca0a [0.895 seconds]
#> • start branch analysis_58_f0ac3217
#> • built branch analysis_58_f0ac3217 [0.848 seconds]
#> • start branch analysis_58_35d814c0
#> ✖ error branch analysis_58_35d814c0
#> • end pipeline [3.56 seconds]
#> Warning messages:
#> 1: NaNs produced 
#> 2: 1 targets produced warnings. Run tar_meta(fields = warnings, complete_only = TRUE) for the messages. 
#> Error:
#> ! Error running targets::tar_make()
#>   Target errors: targets::tar_meta(fields = error, complete_only = TRUE)
#>   Tips: https://books.ropensci.org/targets/debugging.html
#>   Last error: missing values in object
```

Remember, if you just want to see the results that succeeded, run the pipeline with `error = "null"` in `tar_option_set()`. This temporary workaround is especially helpful with so many simulations.

```{r, eval = FALSE}
# _targets.R
library(targets)
library(tarchetypes)

tar_option_set(
  packages = c("broom", "broom.mixed", "dplyr", "nlme", "tibble", "tidyr"),
  error = "null"
)

# Functions etc...
```

```{r, eval = FALSE}
# R console
tar_make()
#> • start target analysis_batch
#> • built target analysis_batch [0.002 seconds]
#> • start branch analysis_58_550d992c
#> • built branch analysis_58_550d992c [1.252 seconds]
#> • start branch analysis_58_582bca0a
#> • built branch analysis_58_582bca0a [0.864 seconds]
#> • start branch analysis_58_f0ac3217
#> • built branch analysis_58_f0ac3217 [0.838 seconds]
#> • start branch analysis_58_35d814c0
#> ✖ error branch analysis_58_35d814c0
#> ✖ error branch analysis_58_35d814c0
#> • start branch analysis_58_e8c6aeab
#> • built branch analysis_58_e8c6aeab [0.774 seconds]
#> # More targets...
#> # ...
#> • start target analysis
#> • built target analysis [0.008 seconds]
#> • end pipeline [27.823 seconds]

# Read the simulations that succeeded.
tar_read(analysis)
#> # A tibble: 270 × 12
#>    term        estimate std.error statistic      p.value conf.low conf.high units tar_batch tar_rep   tar_seed tar_group
#>    <chr>          <dbl>     <dbl>     <dbl>        <dbl>    <dbl>     <dbl> <dbl>     <int>   <int>      <int>     <int>
#>  1 (Intercept)    0.758    0.234       3.24 0.00137        0.300      1.22     58         1       1 -633351515         1
#>  2 factor         0.306    0.0696      4.40 0.0000168      0.170      0.443    58         1       1 -633351515         1
#>  3 (Intercept)    0.524    0.229       2.28 0.0232         0.0745     0.974    58         1       2 -915590912         1
#>  4 factor         0.402    0.0708      5.68 0.0000000402   0.263      0.541    58         1       2 -915590912         1
#>  5 (Intercept)    0.797    0.195       4.08 0.0000624      0.414      1.18     58         1       3 1619222314         1
#>  6 factor         0.295    0.0611      4.82 0.00000257     0.175      0.414    58         1       3 1619222314         1
#>  7 (Intercept)    0.639    0.216       2.96 0.00336        0.216      1.06     58         1       4  825884824         1
#>  8 factor         0.376    0.0723      5.21 0.000000424    0.235      0.518    58         1       4  825884824         1
#>  9 (Intercept)    0.850    0.151       5.63 0.0000000521   0.554      1.15     58         1       5 2080314913         1
#> 10 factor         0.270    0.0507      5.31 0.000000252    0.170      0.369    58         1       5 2080314913         1
#> # … with 260 more rows
#> # ℹ Use `print(n = ...)` to see more rows
```

Now let's seriously debug this pipeline. If each call to `simulate_and_analyze_one_dataset()` takes a long time to run, then the first step is to set one rep per batch in `tar_map_rep()` while keeping the total number of reps the same. In other words, increase `batches` from 20 to 100 and decrease `reps` from 5 to 1.[^seed_batching]. Also remove the `units = 70` scenario because we can reproduce the error without it. 

[^seed_batching]: In `tarchetypes` version `0.7.1.9000` and above, this re-batching will [not change the random number generator seed assigned to each call to `simulate_and_analyze_one_dataset()`](https://github.com/ropensci/tarchetypes/pull/113).

```{r, eval = FALSE}
# _targets.R
# packages, options, and functions...
list(
  tar_map_rep(
    name = analysis,
    command = simulate_and_analyze_one_dataset(units),
    values = data.frame(units = 58), # Remove the units = 70 scenario.
    names = all_of("units"),
    batches = 100, # 100 batches now
    reps = 1 # 1 rep per batch now
  )
)
```

```{r, eval = FALSE}
# R console
tar_make()
#> • start target analysis_batch
#> • built target analysis_batch [0.002 seconds]
#> • start branch analysis_58_550d992c
#> • built branch analysis_58_550d992c [0.502 seconds]
#> • start branch analysis_58_582bca0a
#> • built branch analysis_58_582bca0a [0.208 seconds]
#> • start branch analysis_58_f0ac3217
#> # More successful targets...
#> # ...
#> • start branch analysis_58_b59aa384
#> ✖ error branch analysis_58_b59aa384
#> • end pipeline [3.828 seconds]
#> Warning messages:
#> 1: NaNs produced 
#> 2: 1 targets produced warnings. Run tar_meta(fields = warnings, complete_only = TRUE) for the messages. 
#> Error:
#> ! Error running targets::tar_make()
#>   Target errors: targets::tar_meta(fields = error, complete_only = TRUE)
#>   Tips: https://books.ropensci.org/targets/debugging.html
#>   Last error: missing values in object
```

Around 20 targets ran successfully, and target `analysis_58_b59aa384` hit an error. Let's interactively debug `analysis_58_b59aa384` without interfering with any other targets:

1. Set the `debug` option to `"analysis_58_b59aa384"` in `tar_option_set()`.
1. Restart your R session to remove detritus from memory.
1. Run `tar_make(callr_function = NULL)`.[^interactive]

[^interactive]: You can also run `tar_make_clustermq(callr_function = NULL)` or `tar_make_future(callr_function = NULL)`. In either case, the target to debug will not run on a parallel worker even if you set `deployment = "worker"`.

```{r, eval = FALSE}
# _targets.R
library(targets)
library(tarchetypes)

tar_option_set(
  packages = c("broom", "broom.mixed", "dplyr", "nlme", "tibble", "tidyr"),
  debug = "analysis_58_b59aa384" # Set the target you want to debug.
)

# Functions etc...
```

```{r, eval = FALSE}
# R console
library(targets)

# Restart your R session.
rstudioapi::restartSession()

# Run the pipeline in your interactive R session (no callr process)
tar_make(callr_function = NULL)
#>✔ skip target analysis_batch
#>✔ skip branch analysis_58_550d992c
#>✔ skip branch analysis_58_582bca0a
#>✔ skip branch analysis_58_f0ac3217
#>✔ skip branch analysis_58_35d814c0
#>✔ skip branch analysis_58_e8c6aeab
#>✔ skip branch analysis_58_18f79420
#>✔ skip branch analysis_58_1af19a55
#>✔ skip branch analysis_58_5e604f91
#>✔ skip branch analysis_58_5d2c5812
#>✔ skip branch analysis_58_729b7859
#>✔ skip branch analysis_58_d3899b7b
#>✔ skip branch analysis_58_2a182d3f
#>✔ skip branch analysis_58_5be362d3
#>✔ skip branch analysis_58_5d86137a
#>✔ skip branch analysis_58_a3562efd
#>✔ skip branch analysis_58_a6d57bfd
#>✔ skip branch analysis_58_f15d092c
#>✔ skip branch analysis_58_53efc7f5
#>• start branch analysis_58_b59aa384
#>• pause pipeline
#>  debug target analysis_58_b59aa384
#>
#>ℹ You are now running an interactive debugger.
#>  You can enter code and print objects as with the normal R console.
#>  How to use: https://adv-r.hadley.nz/debugging.html#browser
#>
#>ℹ The debugger is poised to run the command of target analysis_58_b59aa384:
#>
#>     tarchetypes::tar_rep_run(command = tarchetypes::tar_append_static_values(object = simulate_and_analyze_one_dataset(58), 
#>    values = list(units = 58)), batch = analysis_batch, reps = 1, 
#>    iteration = "vector")
#>
#>ℹ Tip: run debug(tarchetypes::tar_rep_run) and then enter "c"
#>  to move the debugger inside function tarchetypes::tar_rep_run().
#>  Then debug the function as you would normally (without {targets}).
#>Called from: eval(expr = expr, envir = envir)
Browse[1]>
```

At this point, we are in an interactive debugger again. Only this time, we quickly skipped straight to the target we want to debug. We can follow the advice in the prompt above, or we can tinker in other ways.

```{r, eval = FALSE}
# R console
# Jump to the function we want to debug.
Browse[1]> debug(analyze_data)
Browse[1]> c # Continue to the next breakpoint.
#> debugging in: analyze_data(data)

# Tinker with the R session to see if you can reproduce the error.
Browse[2]> model <- gls(
+   model = outcome ~ factor,
+   data = data,
+   correlation = corSymm(form = ~ measurement | unit),
+   weights = varIdent(form = ~ 1 | measurement)
+ )
#> Error in na.fail.default(list(measurement = c(1L, 2L, 3L, 4L, 1L, 2L,  : 
#>   missing values in object

# Figure out what it would take to fix the error.
Browse[1]>  model <- gls(
+   model = outcome ~ factor,
+   data = data,
+   correlation = corSymm(form = ~ measurement | unit),
+   weights = varIdent(form = ~ 1 | measurement),
+   na.action = na.omit
+ )

# Confirm that the bug is fixed.
Browse[1]> tidy(model, conf.int = TRUE, conf.level = 0.95)
#> # A tibble: 2 × 7
#>   term        estimate std.error statistic   p.value conf.low conf.high
#> * <chr>          <dbl>     <dbl>     <dbl>     <dbl>    <dbl>     <dbl>
#> 1 (Intercept)    0.925    0.205       4.51 0.0000103    0.523     1.33 
#> 2 factor         0.279    0.0646      4.32 0.0000230    0.153     0.406
```

## Workspaces

A workspace is a special file that helps locally reconstruct the environment of a target outside the pipeline. To demonstrate, consider a different version of the above example which saves the datasets and the models in different targets. We set `workspace_on_error = TRUE` in `tar_option_set()` so that each errored target proactively saves a workspace file.[^workspace_save]

[^workspace_save]: `tar_option_set()` also has a `workspaces` argument to let you choose which targets save workspace files, regardless of whether they hit errors.

```{r, eval = FALSE}
# _targets.R
library(targets)
library(tarchetypes)

tar_option_set(
  packages = c("broom", "broom.mixed", "dplyr", "nlme", "tibble", "tidyr"),
  workspace_on_error = TRUE # Save a workspace file for a target that errors out.
)

simulate_data <- function(units) {
  tibble(unit = seq_len(units), factor = rnorm(units, mean = 3)) %>%
    expand_grid(measurement = seq_len(4)) %>%
    mutate(outcome = sqrt(factor) + rnorm(n()))
}

analyze_data <- function(data) {
  gls(
    model = outcome ~ factor,
    data = data,
    correlation = corSymm(form = ~ measurement | unit),
    weights = varIdent(form = ~ 1 | measurement)
  ) %>%
    tidy(conf.int = TRUE, conf.level = 0.95)
}

list(
  tar_target(rep, seq_len(100)),
  tar_target(data, simulate_data(100), pattern = map(rep)),
  tar_target(analysis, analyze_data(data), pattern = map(data))
)
```


```{r, eval = FALSE}
# R console
tar_make()
#> • start target rep
#> • built target rep [0.6 seconds]
#> • start branch data_c9beb7ca
#> • built branch data_c9beb7ca [0.015 seconds]
#> • start branch data_3658ffe6
#> • built branch data_3658ffe6 [0.009 seconds]
#> • start branch data_b786c6ae
#> # More data targets...
#> # ...
#> • built pattern data
#> • start branch analysis_3ccf0b08
#> • built branch analysis_3ccf0b08 [0.214 seconds]
#> • start branch analysis_0e85e530
#> • built branch analysis_0e85e530 [0.227 seconds]
#> • start branch analysis_03816a9f
#> # More analysis targets...
#> # ...
#> • start branch analysis_02de2921
#> • record workspace analysis_02de2921
#> ✖ error branch analysis_02de2921
#> • end pipeline [5.619 seconds]
#> There were 15 warnings (use warnings() to see them)
#> Error:
#> ! Error running targets::tar_make()
#>   Target errors: targets::tar_meta(fields = error, complete_only = TRUE)
#>   Tips: https://books.ropensci.org/targets/debugging.html
#>   Last error: missing values in object
```

What went wrong with target `analysis_02de2921`? To find out, we load the workspace in an interactive session.

```{r, eval = FALSE}
# R console
# List the available workspaces.
tar_workspaces()
#> [1] "analysis_02de2921"

# Load the workspace.
tar_workspace(analysis_02de2921)
```

At this point, the global objects, functions, and upstream dependencies of target `analysis_02de2921` are in memory. In addition, the target's original random number generator seed is set.[^seed]

[^seed]: You can retrieve this seed with `tar_meta(names = analysis_02de2921, fields = seed)`. In the pipeline, `targets` sets this seed with `withr::with_seed()` just before running the target. However, other functions or target factories may set their own seeds. For example, `tarchetypes::tar_map_rep()` sets its own target seeds so they are [resilient to re-batching](https://github.com/ropensci/tarchetypes/pull/113). For more details on seeds, see the documentation of the `seed` argument of [`tar_option_set()`](https://docs.ropensci.org/targets/reference/tar_option_set.html#arguments).

```{r, eval = FALSE}
# R console
ls()
#> [1] "analyze_data"  "data"          "simulate_data"
```

With the data and functions in hand, you can reproduce the error locally.

```{r, eval = FALSE}
# R console
analyze_data(data)
#> Error in na.fail.default(list(measurement = c(1L, 2L, 3L, 4L, 1L, 2L,  : 
#>   missing values in object
```

With more interactive tinkering your local R session, you can find the bug.

## Getting help

If you are still stuck, you ask questions at <https://github.com/ropensci/targets/discussions>. Please keep the following in mind.

* Before posting, search previous issues and discussions to see if your issue is already solved ([`targets` discussions](https://github.com/ropensci/targets/discussions), [`targets` issues](https://github.com/ropensci/targets/issues), [`tarchetypes` discussions](https://github.com/ropensci/tarchetypes/discussions), and [`tarchetypes` issues](https://github.com/ropensci/tarchetypes/issues)).
* Try to isolate the problem and reproduce it without using `targets`. If `targets` is not the root cause, please post on [Stack Overflow](https://stackoverflow.com) etc. and discuss only the packages that are directly relevant to the problem.
* Share a [minimal, runnable, reproducible example](https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example): minimal so others can understand it easily, and runnable so others can empirically search for the root cause of the problem.
