# Dynamic branching {#dynamic}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = TRUE)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(targets)
library(dplyr)
```

## Branching

Sometimes, a pipeline contains more targets than a user can comfortably type by hand. For projects with hundreds of targets, branching can make the code in `_targets.R` shorter and more concise. 

`targets` supports two types of branching: dynamic branching and [static branching](#static). Some projects are better suited to dynamic branching, while others benefit more from [static branching](#static) or a combination of both. Here is a short list of tradeoffs.

Dynamic | Static
---|---
Pipeline creates new targets at runtime. | All targets defined in advance.
Cryptic target names. | Friendly target names.
Scales to hundreds of branches. | Does not scale as easily for `tar_visnetwork()` etc.
No metaprogramming required. | Familiarity with metaprogramming is helpful.

## About dynamic branching

Dynamic branching is the act of defining new targets (called branches) while the pipeline is running (e.g. during `tar_make()`). Prior to launching the pipeline, the user does not need to know the number of branches or the input data of each branch.

To use dynamic branching, set the `pattern` argument of `tar_target()`. The pattern determines how dynamic branches are created and how the input data is partitioned among the branches. A branch is single iteration of the target's command on a single piece of the input data. Branches are automatically created based on how the input data breaks into pieces, and `targets` automatically combines the output from all the branches when you reference the dynamic target as a whole.

## Example

To illustrate, consider the example pipeline below. It uses dynamic branching to generate random [spirographs](https://wjschne.github.io/spiro/index.html).^[The pipeline uses code borrowed from the [`spiro`](https://wjschne.github.io/spiro/index.html) package. The code in the `spirograph_points()` function is adapted from <https://github.com/wjschne/spiro/blob/87f73ec37ceb0a7a9d09856ada8ae28d587a2ebd/R/spirograph.R>
under the CC0 1.0 Universal license: <https://github.com/wjschne/spiro/blob/87f73ec37ceb0a7a9d09856ada8ae28d587a2ebd/LICENSE.md>]. A spirograph is a type of two-dimensional algebraic curve determined (in part) by parameters `fixed_radius` and `cycling_radius`. Targets `fixed_radius` and `cycling_radius` draw random parameter values, and the dynamic target `points` generates a spirograph dataset for each set of parameters (one spirograph per dynamic branch). Target `single_plot` plots each spirograph separately, and `combine_plot` plots all the spirographs together.

```{r, echo = FALSE, eval = TRUE}
library(targets)
tar_script({
  library(ggplot2)
  library(targets)
  library(tibble)
  
  # From https://github.com/wjschne/spiro/blob/87f73ec37ceb0a7a9d09856ada8ae28d587a2ebd/R/spirograph.R
  # Adapted under the CC0 1.0 Universal license: https://github.com/wjschne/spiro/blob/87f73ec37ceb0a7a9d09856ada8ae28d587a2ebd/LICENSE.md
  spirograph_points <- function(fixed_radius, cycling_radius) {
    t <- seq(1, 30 * pi, length.out = 1e4)
    diff <- (fixed_radius - cycling_radius)
    ratio <- diff / cycling_radius
    x <- diff * cos(t) + cos(t * ratio)
    y <- diff * sin(t) - sin(t * ratio)
    label <- "fixed_radius = %s, cycling_radius = %s"
    parameters <- sprintf(label, fixed_radius, cycling_radius)
    tibble(x = x, y = y, parameters = parameters)
  }
  
  plot_spirographs <- function(points) {
    ggplot(points) +
      geom_point(aes(x = x, y = y, color = parameters), size = 0.1) +
      facet_wrap(~parameters) +
      theme_gray(16) +
      guides(color = "none")
  }
  
  list(
    tar_target(fixed_radius, sample.int(n = 10, size = 2)),
    tar_target(cycling_radius, sample.int(n = 10, size = 2)),
    tar_target(
      points,
      spirograph_points(fixed_radius, cycling_radius),
      pattern = map(fixed_radius, cycling_radius)
    ),
    tar_target(
      single_plot,
      plot_spirographs(points),
      pattern = map(points),
      iteration = "list"
    ),
    tar_target(combined_plot, plot_spirographs(points))
  )
})
```

```{r, eval = FALSE}
# _targets.R
library(ggplot2)
library(targets)
library(tibble)

# From https://github.com/wjschne/spiro/blob/87f73ec37ceb0a7a9d09856ada8ae28d587a2ebd/R/spirograph.R
# Adapted under the CC0 1.0 Universal license: https://github.com/wjschne/spiro/blob/87f73ec37ceb0a7a9d09856ada8ae28d587a2ebd/LICENSE.md
spirograph_points <- function(fixed_radius, cycling_radius) {
  t <- seq(1, 30 * pi, length.out = 1e4)
  diff <- (fixed_radius - cycling_radius)
  ratio <- diff / cycling_radius
  x <- diff * cos(t) + cos(t * ratio)
  y <- diff * sin(t) - sin(t * ratio)
  label <- "fixed_radius = %s, cycling_radius = %s"
  parameters <- sprintf(label, fixed_radius, cycling_radius)
  tibble(x = x, y = y, parameters = parameters)
}

plot_spirographs <- function(points) {
  ggplot(points) +
    geom_point(aes(x = x, y = y, color = parameters), size = 0.1) +
    facet_wrap(~parameters) +
    theme_gray(16) +
    guides(color = "none")
}

list(
  tar_target(fixed_radius, sample.int(n = 10, size = 2)),
  tar_target(cycling_radius, sample.int(n = 10, size = 2)),
  tar_target(
    points,
    spirograph_points(fixed_radius, cycling_radius),
    pattern = map(fixed_radius, cycling_radius)
  ),
  tar_target(
    single_plot,
    plot_spirographs(points),
    pattern = map(points),
    iteration = "list"
  ),
  tar_target(combined_plot, plot_spirographs(points))
)
```

```{r, eval = TRUE}
tar_make()
```

The final plot shows all the spirographs together.

```{r, eval = TRUE}
tar_read(combined_plot)
```

This plot comes from all the branches of `points` aggregated together. Because target `points` has `iteration = "vector"` in `tar_target()`, any reference to the whole target automatically aggregates the branches using `vctrs::vec_c()`. For data frames, this just binds all the rows.

```{r, eval = TRUE}
tar_read(points)
```

By contrast, target `single_plot` list of branches because of `iteration = "list"`.

```{r}
tar_load(single_plot)
class(single_plot)
length(single_plot)
```

Use the `branches` argument of `tar_read()` to read an individual branch or subset of branches.

```{r}
tar_read(single_plot, branches = 1)
```


## Pattern types

`targets` supports many more types of dynamic branching patterns.

* `map()`: one branch per tuple of elements.
* `cross()`: one branch per *combination* of elements.
* `slice()`: select individual pieces to branch over. For example, `pattern = slice(x, index = c(3, 4))` branches over the third and fourth slices (or branches) of target `x`.
* `head()`: branch over the first few elements.
* `tail()`: branch over the last few elements.
* `sample()`: branch over a random subset of elements.

Patterns are composable. For example, `pattern = cross(other_parameter, map(fixed_radius, cycling_radius))` is conceptually equivalent to `tidyr::crossing(other_parameter, tidyr::nesting(fixed_radius, cycling_radius))`. You can test and experiment with branching structures using [`tar_pattern()`](https://docs.ropensci.org/targets/reference/tar_pattern.html). In the output below, suffixes `_1`, `_2`, and `_3`, denote both dynamic branches and the slices of upstream data they branch over.

```{r, eval = TRUE}
tar_pattern(
  cross(other_parameter, map(fixed_radius, cycling_radius)),
  other_parameter = 3,
  fixed_radius = 2,
  cycling_radius = 2
)
```

## Iteration

The `iteration` argument of `tar_target()` determines how to split non-dynamic targets and how to aggregate dynamic ones. Consider the following subset of the spirograph pipeline. Below, `iteration` is equal to `"vector"` by default.

```{r, eval = FALSE}
list(
  tar_target(fixed_radius, sample.int(n = 10, size = 2)),
  tar_target(cycling_radius, sample.int(n = 10, size = 2)),
  tar_target(
    points,
    spirograph_points(fixed_radius, cycling_radius),
    pattern = map(fixed_radius, cycling_radius)
  ),
  tar_target(combined_plot, plot_spirographs(points))
)
```

For non-dynamic targets `fixed_radius` and `cycling_radius`, `iteration = "vector"` means that downstream dynamic targets branch over slices from `vctrs::vec_slice()`. In other words, the first branch of `points` is `spirograph_points(vec_slice(fixed_radius, 1), vec_slice(cycling_radius, 1))`, and second branch of `points` is `spirograph_points(vec_slice(fixed_radius, 2), vec_slice(cycling_radius, 2))`. Since `fixed_radius` is a numeric vector, `vec_slice(fixed_radius, 1)` is a vector of length 1. (But if `fixed_radius` were a data frame, then `vec_slice(fixed_radius, 1)` would be a one-row data frame). If `iteration` were equal to `"list"` for `fixed_radius` and `cycling_radius`, then the first and second branches would instead be `spirograph_points(fixed_radius[[1]], cycling_radius[[1]])` and `spirograph_points(fixed_radius[[2]])`, respectively.

For dynamic target `points`, `iteration = "vector"` means all the branches are aggregated with `vctrs::vec_c()`. Since each branch of `points` is a data frame, aggregation is equivalent to `dplyr::bind_rows()` in this case. That is why, as previously shown, `tar_read(points)` returns a monolithic data frame. `points` becomes a monolithic data frame also when non-dynamic target `combined_plot` runs its command `plot_spirographs(points)`. In other words, `plot_spirographs(points)` is equivalent to `plot_spirographs(vec_c(points_940d58fc, points_bf28fe5a))`, where `points_940d58fc` and `points_bf28fe5a` are the individual branches of `points`.^[`points_940d58fc` and `points_bf28fe5a` are targets in their own right and can be inspected with `tar_read(points_940d58fc)` and `tar_read(points_bf28fe5a)`]. If `iteration` were equal to `"list"` for `points`, then `plot_spirographs(points)` in `combined_plot` would be equivalent to `plot_spirographs(list(points_940d58fc, points_bf28fe5a))`.

`iteration = "group"` is for dynamic branching across `dplyr::group_by()` row groups of a data frame, and it is covered in the next section.

## Branching over row groups



## Dynamic branching performance

With dynamic branching, it is super easy to create an enormous number of targets. But when the number of targets starts to exceed a couple hundred, `tar_make()` slows down, and graphs from `tar_visnetwork()` start to become unmanageable.

If that happens to you, you might be tempted to use the `names` and `shortcut` arguments to `tar_make()`, e.g. `tar_make(names = all_of("only", "these", "targets"), shortcut = TRUE)`, to completely temporarily omit thousands of upstream targets for the sake of concentrating on one section at a time. However, this technique is only a temporary measure, and it is best to eventually revert back to the default `names = NULL` and `shortcut = FALSE` to ensure reproducibility.

Another temporary workaround is to temporarily select subsets of branches. For example, instead of `pattern = map(large_target)` in `tar_target()`, you could prototype on a target that uses `pattern = head(map(large_target), n = 1)` or `pattern = slice(map(large_target), c(4, 5, 6))`. In the case of `slice()`, the `tar_branch_index()` function (only in `targets` version 0.5.0.9000 and above) can help you find the required integer indexes  corresponding to individual branch names you may want.

However, both the above solutions are temporary workarounds for the sake of prototyping a pipeline and chipping away at a large computation. To help the entire pipeline scale more efficiently, consider batching your work into a smaller number of targets.

## Batching

[Targetopia](https://wlandau.github.io/targetopia.html) packages usually have functions that support batching for various use cases. In [`stantargets`](https://wlandau.github.io/stantargets/), [`tar_stan_mcmc_rep_summary()`](https://wlandau.github.io/stantargets/articles/mcmc_rep.html) and friends automatically use batching behind the scenes. The user simply needs to select the number of batches and number of reps per batch. Each batch is a dynamic branch with multiple reps, and each rep fits the user's model once and computes summary statistics. 

In [`tarchetypes`](https://docs.ropensci.org/tarchetypes/), [`tar_rep()`](https://docs.ropensci.org/tarchetypes/reference/tar_rep.html) is a general-use target factory for dynamic branching. It allows you to repeat arbitrary code over multiple reps split into multiple batches. Each batch gets its own reproducible random number seed generated from the target name (as do all targets) and reps run sequentially within each batch, so the results are reproducible. 
The [`targets-stan`](https://github.com/ropensci/targets-stan) repository has an example of batching implemented from scratch. The goal of the pipeline is to validate a Bayesian model by simulating thousands of dataset, analyzing each with a Bayesian model, and assessing the overall accuracy of the inference. Rather than define a target for each dataset in model, the pipeline breaks up the work into batches, where each batch has multiple datasets or multiple analyses. Here is a version of the pipeline with 40 batches and 25 simulation reps per batch (1000 reps total in a pipeline of 82 targets).

```{r, eval = FALSE}
# _targets.R
library(targets)
list(
  tar_target(model_file, compile_model("stan/model.stan"), format = "file"),
  tar_target(index_batch, seq_len(40)),
  tar_target(index_sim, seq_len(25)),
  tar_target(
    data_continuous,
    purrr::map_dfr(index_sim, ~simulate_data_continuous()),
    pattern = map(index_batch)
  ),
  tar_target(
    fit_continuous,
    map_sims(data_continuous, model_file = model_file),
    pattern = map(data_continuous)
  )
)
```
