---
execute:
  freeze: auto
---

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = TRUE)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(targets)
```

# Help {#help}

This chapter explains how to ask for assistance. To ensure a constructive dialogue and a positive outcome, please read the entire page before reaching out.

## Is it really about `targets`?

The `targets` package itself is rarely the cause of an issue in a `targets` pipeline. Most problems come from the user-defined R code, R commands, and R functions that the pipeline calls, as well as other R packages on your system. Before posting, please attempt to troubleshoot and figure out if `targets` is really the cause, using the tips in the [debugging chapter](#debugging) as needed. If it turns out to be a non-`targets` issue, then please ask your question in a non-`targets` forum and rewrite the question accordingly.

## Search existing posts

Please check if your question already has an answer. The bottom of this page links to searchable forums with hundreds of existing discussions. If a relevant thread already exists, please read it first. Then if necessary, please comment there instead of starting a new post.

## Post the complete context

Computer code is extremely strict. In order to run properly, it needs to be exactly correct, down to the last detail. That means for every bug you discover, there are hundreds of strange and unguessable potential causes. The culprit could be a typo buried deep in the `targets` codebase, an [incorrect hidden assumption lost in translation](https://youtu.be/2ve20PVNZ18?t=52) about using `targets`, an obscure bug in an obscure R package dependency, a mistake in one of your R functions, or some other mysterious or unexpected reason. In addition, you may be the first user to experience the problem. So when you ask for help, it is not enough to state the error and ask what might be the solution. The true answer is too surprising to predict, even for the most experienced experts of R and `targets`.

To receive useful feedback, please provide the complete context of the problem, including a reprex. The purpose of a reprex, or reproducible example^[Also known as a minimal reproducible example or minimal working example.], is to eliminate the knowledge gaps and hidden assumptions where bugs hide. A reprex is a sample of complete, self-contained, runnable code that fully emulates and reproduces the problem. The code should [look clean](https://style.tidyverse.org/), be as short and concise as possible, run in as few seconds as possible, and contain only the details most relevant to troubleshooting. You can embed the code inline in your question, or you can upload it to a public repository and post the link. Regardless, please expect that anyone trying to help will read all the code and run the enclosed `_targets.R` file on their own private computer. Please make this empirical process as quick and easy as possible for the people who volunteer their valuable time and energy to respond to you.

The following posts explain how to write a good reprex.

  * <https://stackoverflow.com/help/minimal-reproducible-example>
  * <https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example>
  * <https://reprex.tidyverse.org/>
  * <https://www.tidyverse.org/blog/2017/12/workflow-vs-script/>

The discussion forums of `targets` and `tarchetypes` have excellent reprexes from users.

  * <https://github.com/ropensci/tarchetypes/discussions/90>
  * <https://github.com/ropensci/tarchetypes/discussions/102>
  * <https://github.com/ropensci/tarchetypes/discussions/126>
  * <https://github.com/ropensci/tarchetypes/discussions/129>
  * <https://github.com/ropensci/targets/discussions/884>
  * <https://github.com/ropensci/targets/discussions/953#discussioncomment-3842883>
  * <https://github.com/ropensci/targets/discussions/944>
  * <https://github.com/ropensci/targets/discussions/945>
  * <https://github.com/ropensci/targets/discussions/954>

## Code of conduct

All developers, users, and other discussants must follow the [code of conduct](https://ropensci.org/code-of-conduct/). Please read it before commenting or posting.

## Sensitive information

The forums linked below are open to the public, and anyone with an internet connection can read them. Please do not post any company confidential information, personally identifiable information, private research data, or any other sensitive information.

## R community

You can reach out to the R community at these forums.

* [Stack Overflow](https://stackoverflow.com/)
* [Posit Community](https://community.rstudio.com/)
* [rOpenSci Discuss](https://discuss.ropensci.org)
* [Mastodon](https://mastodon.social) (`#rstats` hashtag)
* [R subreddit](https://www.reddit.com/r/rstats/)

## Contact the maintainer

To contact the maintainer directly, please post to the relevant public [GitHub Discussions](https://github.com/features/discussions) page.^[You will need to create a free [GitHub](https://github.com/) account, which is a straightforward process.] Examples:

* `targets`: <https://github.com/ropensci/targets/discussions>
* `tarchetypes`: <https://github.com/ropensci/tarchetypes/discussions>
* `gittargets`: <https://github.com/ropensci/gittargets/discussions>
* `jagstargets`: <https://github.com/ropensci/jagstargets/discussions>
* `stantargets`: <https://github.com/ropensci/stantargets/discussions>

Anyone can find and link to a public GitHub discussion, and the ability to search and cross-reference helps the users and the maintainer. For this reason, please do not send private emails, instant messages, or mentions on social media.

## Out of office

When the maintainer is out of office, the GitHub repositories above will enter vacation mode. During these times, [Github interactions are temporarily limited](https://docs.github.com/en/communities/moderating-comments-and-conversations/limiting-interactions-in-your-repository), which means users cannot open or comment on [issues](https://docs.github.com/en/issues/tracking-your-work-with-issues/about-issues) or [discussions](https://docs.github.com/en/discussions). Each repository will have an open [issue like this one](https://github.com/ropensci/targets/issues/1007) to explain vacation mode and show the date that the maintainer will return and re-enable interactions.^[I automate these interaction settings and issues using R functions `vacation_on()` and `vacation_off()` from <https://github.com/wlandau/dotfiles/blob/main/github/vacation.R>.]

## Thanks

Thank you for reading this chapter, thank you for taking an active role in troubleshooting, and thank you for respecting the time and effort it takes to help users.
