---
execute:
  freeze: auto
---

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = TRUE)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(targets)
```

# Help {#help}

This chapter explains how to ask for help with specific issues using `targets`. 

## Is it really about `targets`?

The `targets` package itself is rarely the cause of an issue in a `targets` pipeline. Most problems usually come from the user-defined R code, R commands, and R functions that the pipeline calls, as well as other R packages on your system. Please attempt to troubleshoot and figure out if `targets` is really the cause, using the tips in the [debugging chapter](#debugging) as needed. If it turns out to be a non-`targets` issue, then please ask your question in a non-`targets` forum and rewrite the question accordingly.

## Search existing posts

Please check if your question already has an answer. The bottom of this page links to searchable forums with hundreds of existing discussions. If a relevant thread already exists, please read it first. Then if necessary, please comment there instead of starting a new post.

## Post the complete context

Your issue could have any number of unknown causes. It could be a bug in `targets`, a bug in a different R package, a mistake in one of your R functions, or hundreds of other reasons. Because of the endless possibilities, it is not enough to state the error and ask what the cause might be. Not even an expert in R or `targets` can guess the solution, no matter how much prior knowledge or experience they have.

To receive constructive help, please provide the complete context of your use case, including a reprex. A reprex, or reproducible example^[Also known as a minimal reproducible example or minimal working example.], is a piece of complete, self-contained, readable, runnable code that fully emulates the problem. You could embed the code inline in your question, or you could upload it to a public repository and post the link. Regardless, you can expect that anyone trying to help will run the `_targets.R` file in the reprex and tinker with the code you posted (so please make it [easy to read](https://style.tidyverse.org/)). Helping a user troubleshoot is an extremely empirical process.

The following posts explain how to write a good reprex.

  * <https://stackoverflow.com/help/minimal-reproducible-example>
  * <https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example>
  * <https://reprex.tidyverse.org/>
  * <https://www.tidyverse.org/blog/2017/12/workflow-vs-script/>

The `targets` discussion forum has excellent reprexes from users.

  * <https://github.com/ropensci/targets/discussions/884>
  * <https://github.com/ropensci/targets/discussions/953#discussioncomment-3842883>
  * <https://github.com/ropensci/targets/discussions/944>
  * <https://github.com/ropensci/targets/discussions/945>
  * <https://github.com/ropensci/targets/discussions/954>

## Code of conduct

All developers, users, and other discussants must follow the [code of conduct](https://ropensci.org/code-of-conduct/). Please read it before commenting or posting.

## R community

You can reach out to the R community at these forums.

* [Stack Overflow](https://stackoverflow.com/)
* [Posit Community](https://community.rstudio.com/)
* [rOpenSci Discuss](https://discuss.ropensci.org)
* [Mastodon](https://mastodon.social) (`#rstats` hashtag)
* [R subreddit](https://www.reddit.com/r/rstats/)

## Contact the maintainer

To contact the maintainer directly, please post to the relevant public [GitHub Discussions](https://github.com/features/discussions) page.^[You will need to create a free [GitHub](https://github.com/) account, which is a straightforward process.] Examples:

* `targets`: <https://github.com/ropensci/targets/discussions>
* `tarchetypes`: <https://github.com/ropensci/tarchetypes/discussions>
* `gittargets`: <https://github.com/ropensci/gittargets/discussions>
* `jagstargets`: <https://github.com/ropensci/jagstargets/discussions>
* `stantargets`: <https://github.com/ropensci/stantargets/discussions>
