```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = TRUE)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(targets)
```

# Projects {#projects}

A project is a `targets` pipeline together with its supporting source code, data, and configuration settings. This chapter explains best practices when it comes to organizing and configuring `targets` projects.

## Extra reproducibility

For extra reproducibility, it is good practice to use the `renv` R package for package management and Git/Github for code version control. The entire `_targets/` data store should generally not be committed to Git because of its large size.^[However, you may wish to commit `_targets/meta/meta`, which is critical to checking the status of each target and reading targets into memory.] The broader R community has excellent resources and tutorials on getting started with these third-party tools.

## Files

`targets` is mostly indifferent to how you organize the files in your project. However, it is good practice to follow the overall structure of a research compendium or R package (not necessarily with a `DESCRIPTION` file). It also is good practice to give each project its own unique folder with one `targets` pipeline, one `renv` library for package management, and one Git/GitHub repository for code version control. As described later, it is possible to create multiple overlapping projects within a single folder, but this is not recommended for most situations.

The [walkthrough chapter](#walkthrough) shows the file structure for a minimal `targets` project. For more serious projects, the file system may expand to look something like this:

```{r, eval = FALSE}
├── .git/
├── .Rprofile
├── .Renviron
├── renv/
├── index.Rmd
├── _targets/
├── _targets.R
├── _targets.yaml
├── R/
├──── functions_data.R
├──── functions_analysis.R
├──── functions_visualization.R
├── data/
└──── input_data.csv
```

Some of these files are optional, and they have the following roles.

* `.git/`: a folder automatically created by Git for version control purposes.
* `.Rprofile`: a text file automatically created by `renv` to automatically load the project library when you start R at the project root folder. You may wish to add other global configuration here, e.g. declare package precedence using the `conflicted` package.
* `.Renviron`: a text file of key-value pairs defining project-level environment variables, e.g. API keys and package settings. See `Sys.getenv()` for more information on environment variables and how to work with them in R.  
* `index.Rmd`: [Target Markdown](#markdown) report source file to define the pipeline. 
* `_targets/`: the data store where `tar_make()` and similar functions write target storage and metadata when they run the pipeline.
* `_targets.R`: the [target script file](https://docs.ropensci.org/targets/reference/tar_script.html). All `targets` pipelines must have a target script file that returns a target list at the end. If you use [Target Markdown](#markdown) (e.g. `index.Rmd` above) then the target script will be written automatically. Otherwise, you may write it by hand. Unless you apply the custom configuration described later in this chapter, the target script file will always be called `_targets.R` and live at the project root folder.
* `_targets.yaml`: a YAML file to set default arguments to critical functions like `tar_make()`. As described below, you can access and modify this file with functions `tar_config_get()`, `tar_config_set()`, and `tar_config_unset()`. `targets` will attempt to look for `_targets.yaml` unless you set a different path  in the `TAR_CONFIG` environment variable.
* `R/`: directory of scripts containing custom user-defined R code. Most of the code will likely contain [custom functions](#functions) you write to support your targets. You can load these functions with `source("R/function_script.R")` or `eval(parse(text = "R/function_script.R")`, either in a `tar_globals = TRUE` code chunk in [Target Markdown](#markdown) or directly in `_targets.R` if you are not using [Target Markdown](#markdown).
* `data/`: directory of local input data files. As described in the [files chapter](#files), it is good practice to track input files using `format = "file"` in `tar_target()` and then reference those file targets in downstream targets that directly depend on those files.

## Multiple projects

It is generally good practice to give each project its own unique folder with one `targets` pipeline, one `renv` library for package management, and one Git/GitHub repository for code version control. However, sometimes it is reasonable to maintain multiple pipelines within a project: for example, if different pipelines have similar research goals and share the same code base of custom user-defined functions. This section explains how to maintain and navigate such a collection of overlapping projects. 

The functionality below assumes you have `targets` version 0.7.0.9001 or higher, which you may need to install from GitHub.

```{r, eval = FALSE}
remotes::install_github("ropensci/targets")
```

To begin, write the shared code base of custom user-defined functions in `R/`, and write one `targets` pipeline per project. For convenience, we will give each project its own [Target Markdown](#markdown) source file to generate the project's [target script file](https://docs.ropensci.org/targets/reference/tar_script.html), but the same principles generalize to projects without [Target Markdown](#markdown). The file structure looks something like this:

```{r, eval = FALSE}
├── _targets.yaml
├── report_a.Rmd
├── report_b.Rmd
├── R/
...
```

All projects share the same functions defined in the scripts in `R/`, and each project uses a different [target script](https://docs.ropensci.org/targets/reference/tar_script.html) and [data store](#files). To define a different store and script per project, write a top-level `_targets.yaml` configuration to specify these paths explicitly. You can do this from R with `tar_config_set()`.

```{r, eval = FALSE}
tar_config_set(script = "script_a.R", store = "store_a", project = "project_a")
tar_config_set(script = "script_b.R", store = "store_b", project = "project_b")
```

The R code above writes the following `_targets.yaml` configuration file.

```yaml
project_a:
  store: store_a
  script: script_a.R
project_b:
  store: store_b
  script: script_b.R
```

In the setup code chunk of `report_a.Rmd`, set the project to `"project_a"` by setting the `TAR_PROJECT` environment variable.

````
---
title: "Project A"
author: "You"
---

`r ''````{r setup, include = FALSE}
Sys.setenv(TAR_PROJECT = "project_a")
```
````

Likewise for `report_b.Rmd`. 

````
---
title: "Project B"
author: "You"
---

`r ''````{r setup, include = FALSE}
Sys.setenv(TAR_PROJECT = "project_b")
```
````

Then, when you run each report and its `targets` pipeline, [Target Markdown](#markdown) code chunks functions and functions like `tar_make()` will automatically write to the [script](https://docs.ropensci.org/targets/reference/tar_script.html) and [store](#files) defined for each separate project. This is because many default arguments of critical functions come from `tar_config_get()`, which reads from the project-specific configuration in `_targets.yaml`^[To use a different YAML file, set the `TAR_CONFIG` environment variable to the appropriate path.]. The file system will end up looking something like this:

```{r, eval = FALSE}
├── _targets.yaml
├── report_a.Rmd
├── report_b.Rmd
├── script_a.R
├── script_b.R
├── store_a/
├── store_b/
├── R/
...
```

To interactively explore the results of an individual project, set the `TAR_PROJECT` environment variable in your session and then call the appropriate functions.

```{r, eval = FALSE}
Sys.setenv(TAR_PROJECT = "project_a")
tar_visnetwork()
tar_read(your_target)
```

Alternatively, set the script and/or store manually using the arguments of the individual function calls.

```{r, eval = FALSE}
tar_visnetwork(script = "script_a.R", store = "store_a")
tar_read(your_target, store = "store_a")
```
