---
execute:
  freeze: auto
---

# Walkthrough {#walkthrough}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = TRUE)
options(crayon.enabled = FALSE)
```

```{r, include = FALSE, eval = TRUE}
library(fs)
library(targets)
library(tidyverse)
library(withr)
```

```{r, include = FALSE, eval = TRUE}
dir_create("R")
write_csv(airquality, "data.csv")
withr::with_dir("R", {
  tar_script({
    get_data <- function(file) {
      read_csv(file, col_types = cols()) %>%
        filter(!is.na(Ozone))
    }
    fit_model <- function(data) {
      lm(Ozone ~ Temp, data) %>%
        coefficients()
    }
    plot_model <- function(model, data) {
      ggplot(data) +
        geom_point(aes(x = Temp, y = Ozone)) +
        geom_abline(intercept = model[1], slope = model[2])
    }
  })
  file_move("_targets.R", "functions.R")
})
tar_script({
  library(targets)
  source("R/functions.R")
  options(crayon.enabled = FALSE)
  options(tidyverse.quiet = TRUE)
  tar_option_set(packages = c("readr", "dplyr", "ggplot2"))
  list(
    tar_target(file, "data.csv", format = "file"),
    tar_target(data, get_data(file)),
    tar_target(model, fit_model(data)),
    tar_target(plot, plot_model(model, data))
  )
})
```

This chapter walks through a short example of a [`targets`](https://github.com/ropensci/targets)-powered data analysis project. The source code is available at <https://github.com/wlandau/targets-four-minutes>, and you can visit <https://rstudio.cloud/project/3946303> to try out the code in a web browser (no download or installation required). The [documentation website](https://docs.ropensci.org/targets/index.html#examples) links to other examples. The contents of the chapter are also explained in a four-minute video tutorial:

<center>
<iframe src="https://player.vimeo.com/video/700982360?h=38c890bd4f" width="640" height="360" frameborder="0" allow="autoplay; fullscreen; picture-in-picture" allowfullscreen></iframe>
</center>

## About this example

The goal of this short analysis is to assess the relationship among ozone and temperature in base R’s `airquality` dataset. We track a data file, prepare a dataset, fit a model, and plot the model against the data.

## File structure

The file structure of the project looks like this.

```{r, eval = FALSE}
├── _targets.R
├── data.csv
├── R/
│   ├── functions.R
```

`data.csv` contains the data we want to analyze.

```
Ozone,Solar.R,Wind,Temp,Month,Day
36,118,8.0,72,5,2
12,149,12.6,74,5,3
...
```

`R/functions.R` contains our custom user-defined functions. (See the [functions chapter](#functions) for a discussion of function-oriented workflows.)

```{r, eval = FALSE}
# R/functions.R
get_data <- function(file) {
  read_csv(file, col_types = cols()) %>%
    filter(!is.na(Ozone))
}

fit_model <- function(data) {
  lm(Ozone ~ Temp, data) %>%
    coefficients()
}

plot_model <- function(model, data) {
  ggplot(data) +
    geom_point(aes(x = Temp, y = Ozone)) +
    geom_abline(intercept = model[1], slope = model[2])
}
```

## Target script file

Whereas files `data.csv` and `functions.R` are typical user-defined components of a [project-oriented workflow](https://rstats.wtf/project-oriented-workflow.html), the target script file `_targets.R` file is special. Every [`targets`](https://github.com/ropensci/targets) workflow needs a [target script file](https://docs.ropensci.org/targets/reference/tar_script.html) to configure and define the pipeline.^[By default, the target script is a file called `_targets.R` in the project's [root directory](https://martinctc.github.io/blog/rstudio-projects-and-working-directories-a-beginner%27s-guide/). However, you can set the target script file path to something other than `_targets.R`. You can either set the path persistently for your project using [`tar_config_set()`](https://docs.ropensci.org/targets/reference/tar_config_set.html), or you can set it temporarily for an individual function call using the `script` argument of [`tar_make()`](https://docs.ropensci.org/targets/reference/tar_make.html) and related functions.] The [`use_targets()`](https://docs.ropensci.org/targets/reference/use_targets.html) function in `targets` version >= 0.12.0 creates an initial target script for you to fill in. Ours looks like this:

```{r, eval = FALSE}
# _targets.R file
library(targets)
source("R/functions.R")
tar_option_set(packages = c("readr", "dplyr", "ggplot2"))
list(
  tar_target(file, "data.csv", format = "file"),
  tar_target(data, get_data(file)),
  tar_target(model, fit_model(data)),
  tar_target(plot, plot_model(model, data))
)
```

All target script files have these requirements.

1. Load the packages required for configuration, e.g. [`targets`](https://github.com/ropensci/targets) itself.^[target scripts created with `tar_script()` automatically insert a `library(targets)` line at the top by default.]
1. Load your custom functions and small input objects into the R session: in our case, with `source("R/functions.R")`.
1. Use `tar_option_set()` to declare the packages required for analysis, as well as other settings such as the default [storage format](https://docs.ropensci.org/targets/reference/tar_target.html#storage-formats).
1. Write the pipeline at the bottom of `_targets.R`. A pipeline is a list of target objects, which you can create with `tar_target()`. Each target is a step of the analysis. It looks and feels like a variable in R, but it runs reproducibly and stores a value in `_targets/objects/` when we run the pipeline with `tar_make()`.
1. Set additional arguments in [`tar_target()`](https://docs.ropensci.org/targets/reference/tar_target.html#arguments) as needed. For example, `format = "file"` says that the target is an external file, and changes to the contents of the file will invalidate the target (i.e. cause it to rerun).^[`format = "file"` allows you to track multiple files and directories.]

## Inspect the pipeline

Before you run the pipeline for real, it is best to check for obvious errors. [`tar_manifest()`](https://docs.ropensci.org/targets/reference/tar_manifest.html) lists verbose information about each target.

```{r, eval = TRUE}
tar_manifest(fields = all_of("command"))
```

[`tar_visnetwork()`](https://docs.ropensci.org/targets/reference/tar_visnetwork.html) displays the dependency graph. It should show a natural progression of work from left to right as the pipeline progresses, and it targets and functions that depend on one another should have directed edges to identify the dependency relationships.^[If you have hundreds of targets, then [`tar_visnetwork()`](https://docs.ropensci.org/targets/reference/tar_visnetwork.html) may be slow. If that happens, consider temporarily commenting out some targets in `_targets.R` just for visualization purposes.] Dependency relationships are automatically detected using [static code analysis](https://adv-r.hadley.nz/expressions.html#ast-funs), and the order of `tar_target()` calls in the target list does not matter at all.

```{r, eval = TRUE}
tar_visnetwork()
```

## Run the pipeline

`tar_make()` runs the pipeline. It creates a reproducible new external R process which then reads the target script and runs the correct targets in the correct order.^[In `targets` version 0.3.1.9000 and above, you can set the path of the local data store to something other than `_targets/`. A project-level `_targets.yaml` file keeps track of the path. Functions [`tar_config_set()`](https://docs.ropensci.org/targets/reference/tar_config_set.html) and [`tar_config_get()`](https://docs.ropensci.org/targets/reference/tar_config_get.html) can help.]

```{r, eval = TRUE}
tar_make()
```

The output of the pipeline is saved to the `_targets/` data store, and you can read the output with `tar_read()` (see also `tar_load()`).

```{r}
tar_read(plot)
```

The next time you run `tar_make()`, `targets` skips everything that is already up to date, which saves a lot of time in large projects with long runtimes.

```{r, eval = TRUE}
tar_make()
```

You can use `tar_visnetwork()` and `tar_outdated()` to check ahead of time which targets are up to date.

```{r, eval = TRUE}
tar_visnetwork()
```

```{r, eval = TRUE}
tar_outdated()
```

## Changes

The `targets` package notices when you make changes to code and data, and those changes affect which targets rerun and which targets are skipped.^[Internally, special rules called "cues" decide whether a target reruns. The [`tar_cue()`](https://docs.ropensci.org/targets/reference/tar_cue.html) function lets you suppress some of these cues, and the [`tarchetypes`](https://docs.ropensci.org/tarchetypes/) package supports nuanced [cue factories](https://docs.ropensci.org/tarchetypes/reference/index.html#section-cues) and [target factories](https://docs.ropensci.org/tarchetypes/reference/index.html#section-targets-with-custom-invalidation-rules) to further customize target invalidation behavior. The [`tar_cue()`](https://docs.ropensci.org/targets/reference/tar_cue.html) function documentation [explains cues in detail](https://docs.ropensci.org/targets/reference/tar_cue.html#target-invalidation-rules), as well as [specifics on how `targets` detects changes to upstream dependencies](https://docs.ropensci.org/targets/reference/tar_cue.html#dependency-based-invalidation-and-user-defined-functions).]

### Change code

If you change one of your functions, the targets that depend on it will no longer be up to date, and `tar_make()` will rebuild them. For example, let's increase the font size of the plot.

```{r, include = FALSE, eval = TRUE}
withr::with_dir("R", {
  tar_script({
    get_data <- function(file) {
      read_csv(file, col_types = cols()) %>%
        filter(!is.na(Ozone))
    }
    fit_model <- function(data) {
      lm(Ozone ~ Temp, data) %>%
        coefficients()
    }
    plot_model <- function(model, data) {
      ggplot(data) +
        geom_point(aes(x = Temp, y = Ozone)) +
        geom_abline(intercept = model[1], slope = model[2]) +
        theme_gray(24)
    }
  })
  file_move("_targets.R", "functions.R")
})
```

```{r, eval = FALSE}
# Edit functions.R...
plot_model <- function(model, data) {
  ggplot(data) +
    geom_point(aes(x = Temp, y = Ozone)) +
    geom_abline(intercept = model[1], slope = model[2]) +
    theme_gray(24) # Increased the font size.
}
```

`targets` detects the change. `plot` is "outdated" (i.e. invalidated) and the others are still up to date.

```{r, eval = TRUE}
tar_visnetwork()
```

```{r, eval = TRUE}
tar_outdated()
```

Thus, `tar_make()` reruns `plot` and nothing else.^[We would see similar behavior if we changed the R expressions in any `tar_target()` calls in the target script file.]

```{r, eval = TRUE}
tar_make()
```

Sure enough, we have a new plot.

```{r}
tar_read(plot)
```

### Change data

If we change the data file `data.csv`, `targets` notices the change. This is because `file` is a file target (i.e. with `format = "file"` in `tar_target()`), and the return value from last `tar_make()` identified `"data.csv"` as the file to be tracked for changes. Let's try it out. Below, let's use only the first 100 rows of the `airquality` dataset.

```{r, eval = TRUE}
write_csv(head(airquality, n = 100), "data.csv")
```

Sure enough, `raw_data_file` and everything downstream is out of date, so all our targets are outdated.

```{r, eval = TRUE}
tar_visnetwork()
```

```{r, eval = TRUE}
tar_outdated()
```

```{r, eval = TRUE}
tar_make()
```

## Read metadata

:::{.callout-tip}
## Performance

See the [performance chapter](#performance) for options, settings, and other choices to make the pipeline more efficient. This chapter also has guidance for [monitoring the progress of a running pipeline](https://books.ropensci.org/targets/performance.html#monitoring-the-pipeline).
:::
